### 2024-06-10 slides/data.tex
before presentation

\section{Data}

\begin{transitionframe}

    \rmfamily % Font

    \begin{center}
    {\Huge \textbf{\textcolor{white}{Data}}}
    \end{center}

\end{transitionframe}

\begin{frame}

    \frametitle{Data I} % Title
    \framesubtitle{}  % Subtitle
    \rmfamily % Font

    \begin{wideitemize}
        \item \textcolor{fgre}{US Bureau of Labor Statistics (BLS)}  
        \vspace{9pt}
        \begin{wideitemize}
            \item[\textcolor{fgre}{\textbullet}] \textcolor{fgre}{LAUS Database} \(\to\) County-level labor market statistics (1990-2023)
            \item[\textcolor{fgre}{\textbullet}] \textcolor{fgre}{OEWS Database} \(\to\) National industry-specific occupations and wages statistics (1997-2023) % Occupational Employment and Wage Statistics
        \end{wideitemize}
        \item \textcolor{fgre}{Eckert et al. (2020)} \(\to\) County-level industry-specific employment (1975-2016) 
        \item \textcolor{fgre}{USDOL's Changes in Basic Minimum Wages} (1968-2023)
        \item \textcolor{fgre}{US Census Bureau's Population Estimates Program (PEP)} \(\to\) County-level population and demographics (1969-2020)
        \item \textcolor{fgre}{DEA's ARCOS} \(\to\) Opioid distribution data (2006-2014)
    \end{wideitemize}

\end{frame}

\begin{frame}

    \frametitle{Data II} % Title
    \framesubtitle{}  % Subtitle
    \rmfamily % Font

    \begin{wideitemize}
        \item \textcolor{fgre}{Horwitz et al. (2020)} \(\to\) Prescription Drug Monitoring Programs (1990-2019)
        \item \textcolor{fgre}{Centers for Disease Control (CDC)} 
        \vspace{9pt}
        \begin{wideitemize}
            \item[\textcolor{fgre}{\textbullet}] \textcolor{fgre}{National Vital Statistics System} \(\to\) Total drug poisoning mortality by county (2003-2021)
            \item[\textcolor{fgre}{\textbullet}] \textcolor{fgre}{Multiple Cause of Death} \(\to\) Overdose deaths at the state-year level, for different drugs (1999-2020)
        \end{wideitemize}

    \end{wideitemize}

    \vspace{9pt}
    Current working sample:
    \vspace{9pt}

    \begin{wideitemize}
        \item Unemployment, PDMPs, Minimum Wage rates and demographics for counties of 50 States + DC, 1998-2019
        \item County-level industry-weighted wage distribution and death mortality for counties of 50 States + DC, 2003-2016
    \end{wideitemize}

\end{frame}

\begin{frame}

    \frametitle{PDMPs} % Title
    \framesubtitle{}  % Subtitle
    \rmfamily % Font

    Prescription Drug Monitoring Programs (\textcolor{fblu}{PDMPs}): electronic database that tracks controlled substance prescriptions
    \vspace{9pt}

    \begin{wideenumerate}
        \item \textcolor{fblu}{Modern System PDMPs}: the electronic database becomes accessible to any authorized user (eg, physician, pharmacist, or member of law enforcement)
        \item \textcolor{fblu}{Must Query PDMPs}: the law mandates the prescriber to check the database before prescribing a listed opioid
    \end{wideenumerate}

    %\vspace{9pt}
    %2 always follows 1, but 1 does not always lead to 2

\end{frame}